Estimating Access to Rabies Post-Exposure Prophylaxis in Madagascar

### 2018-12-18 accessPEP.Rmd
adding in commune level catchments etc

## Preliminary data processing
```{r directory, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
rm(list = ls())
library(knitr)
library(RImagePalette)
library(jpeg)
library(tidyverse)
library(lubridate)
library(magrittr)
library(rgdal)
library(broom)
knitr::opts_chunk$set(fig.pos = 'h', echo = FALSE, message = FALSE, warning = FALSE, results = 'hide')

@@

## Get color palette from other network fig!
img <- readJPEG("~/Downloads/Dp0j4GNUwAAPPJV.jpeg")
pal <- image_palette(img, 31)
add.alpha <- function(col, alpha=1){
  if(missing(col))
    stop("Please provide a vector of colours.")
  apply(sapply(col, col2rgb)/255, 2, 
                     function(x) 
                       rgb(x[1], x[2], x[3], alpha=alpha))  
}
ctar_metadata$color <- c("#FCC56F","#FFDBE5", "#7A4900", "#CBCDD2", "#0000A6", "#EFF2F1",
                 "#63FFAC", "#B79762", "#004D43", "#8FB0FF", "#997D87", "#FD9C54", "#8362B5",
                 "#578FB0","#5A0007", "#809693", "#D16100", "#1B4400", "#4FC601", "#3B5DFF", 
                 "#4A3B53", "#FF2F80","#61615A", "#BA0900", "#6B7900", "#00C2A0", "#FFAA92",
                 "#FF90C9", "#B903AA", "#FEFFE6", "#E9E9D2")
ctar_metadata$fill <- ctar_metadata$color
ctar_metadata$fill[ctar_metadata$CTAR %in% no_data] <- "#D3D3D3"
ctar_metadata$fill <- add.alpha(ctar_metadata$fill, 0.25)
ctar_metadata$ctar_dist <- mada_district$mdg_dis_co[match(ctar_metadata$District, 
                                                           mada_district$district)]
display_palette(ctar_metadata$color)
dist_catchments %>% left_join(select(ctar_metadata, CTAR, exclude, 
                                     color, fill, ctar_dist, LATITUDE, LONGITUDE)) -> dist_catchments
mada_district@data <- left_join(mada_district@data, dist_catchments)
```

@@

pdf("figs/network.pdf")
plot(mada_district, col = mada_district$fill, border = "grey") 
points(dist_points$dist_long, dist_points$dist_lat, pch = 20, cex = dist_points$size, 
       col = dist_points$color)
segments(dist_lines$LONGITUDE, dist_lines$LATITUDE, x1 = dist_lines$dist_long, y1 = dist_lines$dist_lat, 
         lwd = dist_lines$size*2, col = dist_lines$color)
dev.off()

@@

exp_mat %>% 
  left_join(select(ctar_metadata, CTAR, id_ctar)) %>%
  left_join(select(dist_catchments, mdg_dis_co, CTAR, exclude), by = c("district" = "mdg_dis_co")) %>%
  group_by(district) %>%
  mutate(prop = n/sum(n),
         catch_match = ifelse(CTAR.x == CTAR.y, 1, 0)) -> catch_check
ggplot(data = catch_check, aes(CTAR.x, district)) + 
  geom_tile(aes(fill = prop)) +
  scale_fill_continuous(type = "viridis", name = "proportion") +
  geom_text(label = catch_check$catch_match)
```

@@

@@

## poisson with total pop as offset
mod <- glm(round(bites) ~ ttimes_weighted + offset(log(round(pop))), 
                 data = exps_cleaned, family = poisson)
plot(exps_cleaned$bites, predict(mod, type = "response"))
plot(exps_cleaned$ttimes_weighted, predict(mod, type = "response"))

```{r}
get.burden <- function(covs = covars, pop = covars$pop, model = mod,
                       p_rabid = 0.5, rho_max = 0.9,
                       max_HDR = 25, min_HDR = 5, 
                       dog_rabies_inc = 0.01, human_exp_rate = 0.39, 
                       prob_death = 0.16, nsims = 1000) {
  store_deaths <- store_p_rabid <- matrix(NA, nrow(covs), nsims)

  for (i in 1:nsims) {
    mean_bites <- predict(model, type = "response", newdata = covs)
    bites <- rpois(length(mean_bites), mean_bites)
    mean_exps <- runif(length(pop), min = human_exp_rate*dog_rabies_inc*pop/max_HDR, 
                    max = human_exp_rate*dog_rabies_inc*pop/min_HDR)
    rabid_exps <- rpois(length(mean_bites), mean_exps)
    p_rabid_t <- ifelse(bites*p_rabid/rabid_exps > rho_max, (rho_max*rabid_exps)/bites, p_rabid)
    deaths <- rbinom(length(mean_bites), rabid_exps - round(bites*p_rabid_t), prob_death)
    store_deaths[, i] <- deaths
    store_p_rabid[, i] <- p_rabid_t
  }
  results <- cbind(get.meanCI(store_deaths), get.meanCI(store_p_rabid))
  colnames(results) <- c("deaths_mean", "deaths_lowerCI", "deaths_upperCI",
                         "p_rabid_mean", "p_rabid_lowerCI", "p_rabid_upperCI")
  results <- cbind(covars, results)
  return(results)
}

@@

plot(results_pr20$ttimes_weighted/60, results_pr20$deaths_mean, col = "darkred", pch = 20)
points(results_pr60$ttimes_weighted/60, results_pr60$deaths_mean, col = "red", pch = 1)
segments(x0 = results_pr60$ttimes_weighted/60, x1 = results_pr60$ttimes_weighted/60,
         y0 = results_pr60$deaths_mean, y1 = results_pr20$deaths_mean, col = "red")

results_pr20 <- get.burden(p_rabid = 0.20, rho_max = 0.90)
results_pr60 <- get.burden(p_rabid = 0.20, rho_max = 0.99)

plot(results_pr20$ttimes_weighted/60, results_pr20$deaths_mean, col = "darkred", pch = 20)
points(results_pr60$ttimes_weighted/60, results_pr60$deaths_mean, col = "red", pch = 1)
segments(x0 = results_pr60$ttimes_weighted/60, x1 = results_pr60$ttimes_weighted/60,
         y0 = results_pr60$deaths_mean, y1 = results_pr20$deaths_mean, col = "red")

results_pr20 <- get.burden(p_rabid = 0.60, rho_max = 0.90)
results_pr60 <- get.burden(p_rabid = 0.60, rho_max = 0.99)

plot(results_pr20$ttimes_weighted/60, results_pr20$deaths_mean, col = "darkred", pch = 20)
points(results_pr60$ttimes_weighted/60, results_pr60$deaths_mean, col = "red", pch = 1)
segments(x0 = results_pr60$ttimes_weighted/60, x1 = results_pr60$ttimes_weighted/60,
         y0 = results_pr60$deaths_mean, y1 = results_pr20$deaths_mean, col = "red")
```

```{r}
exps_cleaned %$% plot(ttimes_weighted, bites)

exps_cleaned %>%
  filter(exclude == 0) %$% 
  plot(ttimes_weighted, incidence, bty = 
                        "n", pch = 20)

@@

geom_point(data = ctar_metadata, aes (x = LONGITUDE, y = LATITUDE))
ggplot() +
  geom_polygon(data = mada_dist_gg, aes(x = long, y = lat, group = group, fill = ttimes_weighted)) +
  scale_fill_continuous(type = "viridis") +
  geom_point(data = ctar_metadata, aes (x = LONGITUDE, y = LATITUDE))

@@

>>>>

## Preliminary data processing
```{r directory, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
rm(list = ls())
library(knitr)
library(RImagePalette)
library(jpeg)
library(tidyverse)
library(lubridate)
library(magrittr)
library(rgdal)
library(broom)
select <- dplyr::select
knitr::opts_chunk$set(fig.pos = 'h', echo = FALSE, message = FALSE, warning = FALSE, results = 'hide')

@@

## Get color palette from other network fig!
img <- readJPEG("~/Downloads/Dp0j4GNUwAAPPJV.jpeg")
pal <- image_palette(img, 31)
add.alpha <- function(col, alpha=1){
  if(missing(col))
    stop("Please provide a vector of colours.")
  apply(sapply(col, col2rgb)/255, 2, 
                     function(x) 
                       rgb(x[1], x[2], x[3], alpha=alpha))  
}
ctar_metadata$color <- c("#FCC56F","#FFDBE5", "#7A4900", "#CBCDD2", "#0000A6", "#EFF2F1",
                 "#63FFAC", "#B79762", "#004D43", "#8FB0FF", "#997D87", "#FD9C54", "#8362B5",
                 "#578FB0","#5A0007", "#809693", "#D16100", "#1B4400", "#4FC601", "#3B5DFF", 
                 "#4A3B53", "#FF2F80","#61615A", "#BA0900", "#6B7900", "#00C2A0", "#FFAA92",
                 "#FF90C9", "#B903AA", "#FEFFE6", "#E9E9D2")
ctar_metadata$fill <- ctar_metadata$color
ctar_metadata$fill[ctar_metadata$CTAR %in% no_data] <- "#D3D3D3"
ctar_metadata$fill <- add.alpha(ctar_metadata$fill, 0.25)
## Getting ctar district and commune
ctar_metadata$ctar_dist <- mada_district$mdg_dis_co[match(ctar_metadata$District, 
                                                           mada_district$district)]
pts <- SpatialPoints(cbind(ctar_metadata$LONGITUDE, ctar_metadata$LATITUDE), 
                   proj4string = CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))
ctar_metadata$ctar_comm <- over(pts, mada_communes)$mdg_cm_

display_palette(ctar_metadata$color)
dist_catchments %>% left_join(select(ctar_metadata, CTAR, exclude, 
                                     color, fill, ctar_dist, LATITUDE, LONGITUDE)) -> dist_catchments
mada_district@data <- left_join(mada_district@data, dist_catchments)
dist_catchments$ctar_in <- ifelse(dist_catchments$mdg_dis_co %in% ctar_metadata$ctar_dist, 1, 0)
```

@@

pdf("figs/network_1.pdf")
plot(mada_district, col = mada_district$fill, border = "grey") 
points(jitter(dist_lines$dist_long, factor = 2), jitter(dist_lines$dist_lat, factor = 2)
       , pch = 1, cex = dist_points$size, 
       col = dist_lines$color)
segments(dist_lines$LONGITUDE, dist_lines$LATITUDE, x1 = dist_lines$dist_long, y1 = dist_lines$dist_lat,
         lwd = dist_lines$size*2, col = dist_lines$color)
dev.off()

@@

exp_mat %>% 
  left_join(select(ctar_metadata, CTAR, id_ctar)) %>%
  left_join(select(dist_catchments, mdg_dis_co, CTAR, exclude), by = c("district" = "mdg_dis_co")) %>%
  group_by(district) %>%
  mutate(prop = n/sum(n),
         catch_match = ifelse(CTAR.x == CTAR.y, 1, 0)) -> catch_check
hist(catch_check$prop)
ggplot(data = catch_check, aes(CTAR.x, district)) + 
  geom_tile(aes(fill = prop)) +
  scale_fill_continuous(type = "viridis", name = "proportion") +
  geom_text(label = catch_check$catch_match)

@@

patient_ts %>%
  filter(id_ctar == 1) %$%
  hist(n, breaks = 100, col = "blue", 
       xlab = "Patients reporting", main = "Soanierana Ivongo")
patient_ts %>%
  filter(ctar == "Toamasina") %$%
  hist(n, breaks = 100, 
       xlab = "Patients reporting", main = "Toamasina", col = "red")

@@

## poisson with total pop as offset
mod <- glm(round(bites) ~ ttimes_weighted + offset(log(round(pop))), 
                 data = exps_cleaned, family = poisson)
par(mfrow = c(2, 1), mar = c(4, 4, 1, 1))
plot(exps_cleaned$bites, predict(mod, type = "response"), bty = "l", pch = 20,
     xlab = "Observed bites", ylab = "Predicted bites")
abline(a = 0, b = 1, lty = 2, col = "blue")
legend("topright", col = "blue", legend = "1:1 line", lty = 2, bty = "n", seg.len = 1, 
       inset = c(0.0, -0.1))
plot(exps_cleaned$ttimes_weighted/60, predict(mod, type = "response"), bty = "l", pch = 20,
     xlab = "Travel times (hrs)", ylab = "Bites", col = alpha("blue", 0.7), 
     ylim = c(0, max(exps_cleaned$bites)))
points(exps_cleaned$ttimes_weighted/60, exps_cleaned$bites, bty = "l", pch = 20,
     xlab = "Travel times (hrs)", col = alpha("darkred", 0.7))
legend("topright", col = c(alpha("blue", 0.7), alpha("darkred", 0.7)), legend = c("Predicted", "Observed"), 
       pch = 20, bty = "n")
```

```{r functions for getting burden}
get.burden <- function(covs = covars, pop = covars$pop, model = mod,
                       p_rabid = 0.5, rho_max = 0.9,
                       max_HDR = 25, min_HDR = 5, 
                       dog_rabies_inc = 0.01, human_exp_rate = 0.39, 
                       prob_death = 0.16, nsims = 1000) {
  store_deaths <- store_p_rabid <- matrix(NA, nrow(covs), nsims)

  for (i in 1:nsims) {
    mean_bites <- predict(model, type = "response", newdata = covs)
    bites <- rpois(length(mean_bites), mean_bites)
    mean_exps <- runif(length(pop), min = human_exp_rate*dog_rabies_inc*pop/max_HDR, 
                    max = human_exp_rate*dog_rabies_inc*pop/min_HDR)
    rabid_exps <- rpois(length(mean_bites), mean_exps)
    p_rabid_t <- ifelse(bites*p_rabid/rabid_exps > rho_max, (rho_max*rabid_exps)/bites, p_rabid)
    deaths <- rbinom(length(mean_bites), rabid_exps - round(bites*p_rabid_t), prob_death)
    store_deaths[, i] <- deaths
    store_p_rabid[, i] <- p_rabid_t
  }

  results <- cbind(get.meanCI(store_deaths), get.meanCI(store_p_rabid))
  colnames(results) <- c("deaths_mean", "deaths_lowerCI", "deaths_upperCI",
                         "p_rabid_mean", "p_rabid_lowerCI", "p_rabid_upperCI")
  results <- cbind(covars, results)
  return(results)
}

@@

plot(results_pr20$ttimes_weighted/60, results_pr20$deaths_mean, col = "darkred", pch = 20)
points(results_pr60$ttimes_weighted/60, results_pr60$deaths_mean, col = "red", pch = 20)
segments(x0 = results_pr60$ttimes_weighted/60, x1 = results_pr60$ttimes_weighted/60,
         y0 = results_pr60$deaths_mean, y1 = results_pr20$deaths_mean, col = "red")

results_base <- get.burden(p_rabid = 0.40, rho_max = 0.98)

p_rabid_check <- seq(0.10, 0.80, by = 0.1)
check <- data.frame(p_rabid = 0, district = NA, ttimes = 0, deaths_mean = 0)
for (i in 1:length(p_rabid_check)) {
  print(paste(i, "/", length(p_rabid_check)))
  results <- get.burden(p_rabid = p_rabid_check[i], rho_max = 0.98)
  p_rabid <- rep(p_rabid_check[i], length(results$deaths_mean))
  check_one <- cbind(p_rabid, results$district, results$ttimes, results$deaths_mean)
  colnames(check_one) <- c("p_rabid", "district", "ttimes", "deaths_mean")
  check <- rbind(check, check_one)
}

check <- check[-1, ]
check$base <- rep(results_base$deaths_mean, length(p_rabid_check))
ggplot(data = check, aes(x = as.factor(p_rabid), y = as.factor(reorder(district, ttimes)))) + 
  geom_tile(aes(fill = deaths_mean/base)) +
  scale_fill_gradient2(high = "red", mid = "white", low = "blue", midpoint = 1, name = "Mean deaths")

rho_max_check <- seq(0.90, 0.99, by = 0.01)
check <- data.frame(rho_max = 0, district = NA, ttimes = 0, deaths_mean = 0)
for (i in 1:length(rho_max_check)) {
  print(paste(i, "/", length(rho_max_check)))
  results <- get.burden(p_rabid = 0.4, rho_max = rho_max_check[i])
  rho_max <- rep(rho_max_check[i], length(results$deaths_mean))
  check_one <- cbind(rho_max, results$district, results$ttimes, results$deaths_mean)
  colnames(check_one) <- c("rho_max", "district", "ttimes", "deaths_mean")
  check <- rbind(check, check_one)
}

check <- check[-1, ]
check$base <- rep(results_base$deaths_mean, length(rho_max_check))
check %>%
  group_by(district) %>%
  mutate(min = min(deaths_mean), max = max(deaths_mean), 
         standardized = (deaths_mean - base)/(max - base)) -> check

ggplot(data = check, aes(x = as.factor(p_rabid), y = as.factor(reorder(district, ttimes)))) + 
  geom_tile(aes(fill = standardized)) +
  scale_fill_gradient2(high = "red", mid = "white", low = "blue", midpoint = 0, name = "Mean deaths") 
ggplot(data = check, aes(x = as.factor(p_rabid), y = as.factor(reorder(district, ttimes)))) + 
  geom_tile(aes(fill = deaths_mean)) +
  scale_fill_gradient2(high = "red", mid = "white", low = "blue", midpoint = 0, name = "Mean deaths")

@@

@@

## Commune level mods
```{r commune matching, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
commune_matching <- read.csv("data/commune_check_CH.csv")
mada_communes <- readOGR("output/communes/communes_extracted.shp")
mada_communes$match <- tolower(str_replace_all(mada_communes$commune,"[^[:graph:]]", "/"))

mada_communes@data %>%
  select(match, mtch_ds, mdg_cm_) %>%
  left_join(commune_matching, .) %>%
  select(name, match, district, mdg_cm_) -> comm_matches

ctar_data$name <- tolower(str_replace_all(ctar_data$commune,"[^[:graph:]]", "/"))
left_join(ctar_data, comm_matches) %>%
  distinct(record_id, .keep_all = TRUE) -> ctar_data

## Looking at proportion matched
ctar_data %>%
  group_by(id_ctar) %>%
  summarise(n = n(), unmatched = sum(is.na(mdg_cm_)),
            prop_matched = unmatched/n) -> ctar_data_matched

ctar_metadata <- merge(ctar_metadata, ctar_data_matched)
ctar_metadata <- ctar_metadata[!is.na(ctar_metadata$id_ctar), ]
ggplot(ctar_metadata, aes(CTAR, prop_matched)) +
  geom_point() +
  coord_flip()  +
  ylab(label = "Proportion of patients matched to commune")
```

```{r }
## Ctar data at commune level
ctar_data %>% 
  filter(excl != 1) %>% # exclude contacts
  mutate_at(vars(starts_with("date")), funs(ymd(.))) %>% # format dates
  group_by(year = year(date_de_consultation), commune = mdg_cm_, id_ctar) %>% # group by year and district
  summarize(n = n()) -> exposures # get count of exposures

comm_catchments <- read.csv("output/catchments_commune_unmasked.csv", row.names = 1)
names(comm_catchments) <- c("mdg_cm_", "CTAR", "ttimes_weighted")
comm_catchments %>% left_join(select(ctar_metadata, CTAR, exclude, 
                                     color, fill, ctar_dist, LATITUDE, LONGITUDE)) -> comm_catchments
mada_communes@data <- left_join(mada_communes@data, comm_catchments)

mada_communes@data %>%
select(commune = mdg_cm_, pop = MDG__201, ttimes_weighted, CTAR, exclude) -> covars
mada_communes$ctar_in <- ifelse(mada_communes$mdg_cm_ %in% ctar_metadata$ctar_comm, 1, 0)

## Catchments with no data
no_data <- c("IPM", "Fianarantsoa", "Ambatomainty", "Ambovombe Androy",
             "Antsiranana", "Mandritsara", "Marolambo", "Nosy be", "Sainte Marie",
             "Taolagnaro", "Tsiroanomandidy", "Vangaindrano")
covars$exclude <- 0
covars$exclude[covars$CTAR %in% no_data] <- 1

exposures %>%
  rename(date = year) %>%
  left_join(covars) %>% # add pop + ttimes + reporting
  left_join(clinic_reporting) %>%   # add reporting by clinic + year
  filter(exclude == 0, prop > 0.25) %>%
  group_by(commune, date) %>%
  mutate(bites = n/prop) %>%
  summarise(bites = sum(bites, na.rm = TRUE)) -> exps_cleaned 
exps_cleaned %>%
  group_by(commune) %>%
  summarise(bites = mean(bites)) %>%
  left_join(covars) -> exps_cleaned

## poisson with total pop as offset
exps_cleaned <- filter(exps_cleaned, ttimes_weighted != "Inf")
mod <- glm(round(bites) ~ ttimes_weighted + offset(log(round(pop))), 
                 data = exps_cleaned, family = poisson)

par(mfrow = c(2, 1), mar = c(4, 4, 1, 1))
plot(exps_cleaned$bites, predict(mod, type = "response"), bty = "l", pch = 20,
     xlab = "Observed bites", ylab = "Predicted bites")
abline(a = 0, b = 1, lty = 2, col = "blue")
legend("topright", col = "blue", legend = "1:1 line", lty = 2, bty = "n", seg.len = 1, 
       inset = c(0.0, -0.1))
plot(exps_cleaned$ttimes_weighted/60, predict(mod, type = "response"), bty = "l", pch = 20,
     xlab = "Travel times (hrs)", ylab = "Bites", col = alpha("blue", 0.7), 
     ylim = c(0, max(exps_cleaned$bites)))
points(exps_cleaned$ttimes_weighted/60, exps_cleaned$bites, bty = "l", pch = 20,
     xlab = "Travel times (hrs)", col = alpha("darkred", 0.7))
legend("topright", col = c(alpha("blue", 0.7), alpha("darkred", 0.7)), legend = c("Predicted", "Observed"), 
       pch = 20, bty = "n")

ctar_data %>% 
  filter(excl != 1) %>% # exclude contacts
  mutate_at(vars(starts_with("date")), funs(ymd(.))) %>% # format dates
  group_by(commune = mdg_cm_, id_ctar) %>% # group by year and district
  summarize(n = n()) ->  exp_mat # get count of exposures

comm_catchments$exclude <- 0
comm_catchments$exclude[comm_catchments$CTAR %in% no_data] <- 1

exp_mat %>% 
  left_join(select(ctar_metadata, CTAR, id_ctar)) %>%
  left_join(select(comm_catchments, mdg_cm_, CTAR, exclude), by = c("commune" = "mdg_cm_")) %>%
  group_by(commune) %>%
  mutate(prop = n/sum(n),
         catch_match = ifelse(CTAR.x == CTAR.y, 1, 0)) -> catch_check
hist(catch_check$prop[catch_check$exclude == 0])
hist(catch_check$prop[catch_check$exclude == 1])

## getting var for ctar in
pts <- SpatialPoints(cbind(ctar_metadata$LONGITUDE, ctar_metadata$LATITUDE), 
                   proj4string = CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))
over(pts, mada_communes)$mtch_ds
```
